Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-06-26 analysis/paper/paper.Rmd
paragraphs for kernel plots and hypo-testing

```{r point-pattern-data-setup}
# get centroid and make a list of three periods
ornaments_period_join_cen_all <-
  ornaments_period_join %>% 
  mutate(lon = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[1]])),
         lat = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[2]])))

@@

```{r}
# transform data to PPP format and sp SpatialPoints
library(spatstat)
library(maptools)
library(raster)

@@

```{r plot-all-densities}
par(mfrow = c(1, 3))
# apply function to each data frame in the list
titles <- names(ornaments_period_join_cen)

@@

@@

```{r}

@@

To assess whether the distribution of ornaments present different density that can indicate the changes caused by global trade network brought by the Europeans, point pattern analysis is used to measure and test the density of ornaments across three time periods. We assume a uniform point density across the research area before European contact period, that is, randomly distributed, and test the null hypothesis that is there is no cluster of ornaments across the sapce during the European contact periods.

```{r-gganimate, eval= FALSE}
# test gganimate for measurement of beads by period
# devtools::install_github('thomasp85/gganimate')
library(gganimate)

anim_gold_bead <-
  ggplot(ornaments_intact_gold,
       aes(Length, Width)) +
  geom_point() +
  transition_states(period, 3, 1) +
  labs(title = 'Period: {closest_state}',
       x ='Length (mm) of golden bead',
       y ='Width (mm) of golden bead')

# does not work
anim_gold_bead_spatial <-
  ggplot(ornaments_period_join_gold) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
  transition_states(period, 3, 1) +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))

>>>>

```{r point-pattern-data-tidy}
# get centroid and make a list of three periods
ornaments_period_join_cen_all <-
  ornaments_period_join %>% 
  mutate(lon = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[1]])),
         lat = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[2]])))

@@

```{r transformation-sptial-data}
# transform data to PPP format and sp SpatialPoints
library(spatstat)
library(maptools)
library(raster)

@@

```{r plot-all-densities, fig.cap= "Kernal density map"}
par(mfrow = c(1, 3))
# apply function to each data frame in the list
titles <- names(ornaments_period_join_cen)

@@

Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period. The location where ornaments from were square-based recorded, thus we took the intensity approach that focuses on the average density of points across space to explore their distribution. We randomly assigned points, here the location of ornaments, for each square according to the data, subset three groups for three time periods, and compute the kernal density for each time period for comparison. Kernal density estimations (KDE) can estimate the probability of the density of events across space by creating a continuous, smooth density surface across space. Here we use KDE to visualize the event of the presence of ornaments that were represents by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier.etal]. Density value is assigned to each cell, of which the unit is meter. The results \@ref(fig:plot-all-densities) show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period. Those maps present three consistent sub-region with shifting core area over time. The distribution might indicate different social groups who own more ornaments, and the multiple groups during European contact period might reflect a more corporate mode of social organization. However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create obvious core area. Next step is to test whether the distribution of ornaments is randomly distributed to understand the underlying social indication related to ornaments.

@@

```{r hypo-test-distrubution, fig.cap= "Histogram of simulated ANN values"}

@@

Here we use Monte Carlo method to test for the complete spatial randomness of spatial events based on the average nearest-neighbor distance method. We hypothesize that the distribution of ornaments is consistent with a completely random process, and then simulate the presence of ornaments across the space for one thousand times for each time periods. We compare our data to the random pattern generated by simulated processes to see if the our data is randomly distributed or not. The results \@ref(fig:hypo-test-distrubution) are represented by histogram showing one thousand simulation with the blue line indicating our observed value. The results show that 100% of the simulated values are much more greater than our observed ANN value during the European contact period, which means the ornaments are clustered distributed. Similar result are also observed during the pre-European period but less extreme. The distribution of ornaments was random during the Chinese contact period and about one third of the simulated values are greater than our observed ANN value. This testing explains that the distribution of ornaments during the pre-European and European contact period might reflect the presence of different social groups. Moreover, the cluster of ornaments way more significant during the European contact period that might indicate some extent of control of ornaments.
